The Rockerverse: Packages and Applications for Containerisation with R

### 2020-03-02 manuscript.Rmd

@@

In **high-performance computing**, one use for containers is to run workflows on shared local hardware where teams manage their own high-performance servers.
This can follow one of several design patterns:
Users may deploy containers to hardware as a work environment for a specific project, containers may provide per-user persistent environments, or a single container can act as a common multi-user environment for a server.
In all cases, though, the containerised approach provides several advantages:
First, users may use the same image and thus work environment on desktop and laptop computers.
The first to patterns provide modularity, while the last approach is most similar to a simple shared server.
Second, software updates can be achieved by updating and redeploying the container rather than by tracking local installs on each server.
Third, the containerised environment can be quickly deployed to other hardware, cloud or local, if more resources are neccessary or in case of server destruction or failure.
In any of these cases, users need a method to interact with the containers, be it an IDE exposed over an HTTP port or command-line access via tools such as SSH.
A suitable method must be added to the container recipes.
The Rocker&nbsp;Project provides containers pre-installed with the RStudio&nbsp;IDE.
In cases where users store nontrivial amounts of data for their projects, the data needs to persist beyond the life of the container.
This may be in shared disks, attached network volumes, or in separate storage where it is uploaded between sessions.
In the case of shared disks or network-attached volumes, care must be taken to match user permissions, and of course backups are still neccessary.
When working with multiple servers, an automation framework such as [Ansible](https://www.ansible.com) may be useful for managing users, permisions, and disks along with containers.

@@

[CyVerse](https://cyverse.org) is an open source, NSF-funded cyberinfrastructure platform for the life sciences \citep{merchant2016iplant}.
CyVerse supports a browser-based 'data science workbench' called the [Discovery Environment (DE)](https://cyverse-de.github.io/), which uses a combination of [HTCondor](https://research.cs.wisc.edu/htcondor/) and Kubernetes for orchestrating container based analysis, it integrates with external HPC (High Performance Computing), NSF-XSEDE through [TAPIS] (https://www.tacc.utexas.edu/tapis) API's.
In addition, CyVerse provides a multi-petabyte Data Store based on [iRODS](https://irods.org/, this allows users to share and manage hundreds of GB's of data.
The DE uses an [OAuth](https://oauth.net/2/) system for single sign on (SSO) for its 70+ thousand registered users.
Rocker integration in the DE is designed to provide life science researchers with scalable, compute-intensive R analysis capabilities needed for large and complex datasets (e.g., genomics/multi-omics, GWAS, phenotypic data, spatial data etc.).
The DE runs Docker containers on demand, with users able to integrate containers from DockerHub or other registries \citep{devisetty2016bringing}.
These capabilities give users flexibility similar to Binder, but allows containers to run on larger computational resources (RAM, CPU, Disk), and for longer periods of time (days to weeks).
The Rocker Project's RStudio and Shiny containers are integrated into the DE by deriving new images from Rocker images.
These images include a reverse proxy using `nginx` to handle communication with CyVerse authentication systems.
All running instances are protected through this authentication system and users can invite other registered users to securely access the same instance \citep{rstudio_proxy_2020}.
Users can reference the same base image and add packages to further customize their R environment.
The Rocker base image is cached on all nodes in the Kubernetes environment to speed up launching.
CyVerse maintains its own [GitHub Organization](https://github.com/cyverse-vice/) with Dockerfiles and configuration scripts auto-built on DockerHub.
For additional reproducibility, CyVerse operates a private Docker image registry where a user's published, i.e., publicly available, containers are held.
CyVerse user accounts are free (with some limitations for non-US users) and the documentation repository [CyVerse Learning Center](https://learning.cyverse.org/) allows community members to document their containers to support training and education.

@@

Nonetheless, at least on the level of R packages some **consolidation** seems in order, e.g., to reduce the number of packages creating `Dockerfile`s from R code or controlling the Docker daemon with R code.
It remains to be seen which approach to control Docker, via the Docker API as \pkg{stevedore} or via system calls as \pkg{dockyard}/\pkg{docker}/\pkg{dockr}, is more sustainable, or whether the question will be answered by the endurance of maintainers and sufficient funding.
Similarly, the capturing of environments and their serialisation in form of a `Dockerfile` currently happens at different levels of abstraction, and re-use of functionality seems reasonable, e.g., \pkg{liftr} could generate the environment with \pkg{containerit}, which in turn may use \pkg{dockerfiler} for low-level R objects representing a `Dockerfile` and its instructions.
In this consolidation, the Rocker&nbsp;Project could play the role of a coordinating entity.
Nonetheless, for the moment, it seems that Rocker Project will be used more experimentation and feature growth, e.g., images for GPU-based computing and artificial intelligence.
Even with coding being more and more accepted as a required and achievable skill, an easier access, for example by exposing containerisation benefits via simple user interfaces in the users' IDE, could be an important next step, since currently containerisation happens more in the background for UI-based development (e.g., a `rocker/rstudio` image in the cloud).
Furthermore, a maturing of the Rockerverse packages for managing containers may lead to them being adopted in situations where manual coding is currently required, e.g. in the case of \CRANpkg{RSelentium} or \CRANpkg{drake} (see Sections&nbsp;\nameref{rselenium} and \nameref{drake} respectively).
In specific cases, e.g., for \CRANpkg{analogsea}, the interaction with the Docker daemon may remain too specific to re-use first-order packages to control Docker.

>>>>

\label{binder}

@@

In **high-performance computing**, one use for containers is to run workflows on shared local hardware where teams manage their own high-performance servers.
This can follow one of several design patterns:
Users may deploy containers to hardware as a work environment for a specific project, containers may provide per-user persistent environments, or a single container can act as a common multi-user environment for a server.
In all cases, though, the containerised approach provides several advantages:
First, users may use the same image and thus work environment on desktop and laptop computers.
The first to patterns provide modularity, while the last approach is most similar to a simple shared server.
Second, software updates can be achieved by updating and redeploying the container rather than by tracking local installs on each server.
Third, the containerised environment can be quickly deployed to other hardware, cloud or local, if more resources are neccessary or in case of server destruction or failure.
In any of these cases, users need a method to interact with the containers, be it an IDE exposed over an HTTP port or command-line access via tools such as SSH.
A suitable method must be added to the container recipes.
The Rocker&nbsp;Project provides containers pre-installed with the RStudio&nbsp;IDE.
In cases where users store nontrivial amounts of data for their projects, the data needs to persist beyond the life of the container.
This may be in shared disks, attached network volumes, or in separate storage where it is uploaded between sessions.
In the case of shared disks or network-attached volumes, care must be taken to match user permissions, and of course backups are still neccessary.

[CyVerse](https://cyverse.org) is an open source, NSF-funded cyberinfrastructure platform for the life sciences providing easy access to computing and storage resources \citep{merchant_iplant_2016}.
CyVerse's a browser-based 'data science workbench' is called [Discovery Environment](https://cyverse-de.github.io/) (DE).
DE uses a combination of [HTCondor](https://research.cs.wisc.edu/htcondor/) and Kubernetes for orchestrating container-based analysis and integrates with external HPC, i.e., [NSF-XSEDE](https://www.xsede.org/), through [TAPIS](https://www.tacc.utexas.edu/tapis) API's.
In addition, CyVerse provides a multi-petabyte Data Store based on [iRODS](https://irods.org/) with shared access by users.
The DE runs Docker containers on demand, with users able to integrate bespoke containers from DockerHub or other registries \citep{devisetty_bringing_2016}.
Rocker integration in the DE is designed to provide life science researchers with scalable, compute-intensive R analysis capabilities needed for large and complex datasets (e.g., genomics/multi-omics, GWAS, phenotypic data, spatial data etc.).
These capabilities give users flexibility similar to Binder, but allows containers to run on larger computational resources (RAM, CPU, Disk), and for longer periods of time (days to weeks).
The Rocker Project's RStudio and Shiny containers are integrated into the DE by deriving new images from Rocker images\footnote{See \href{https://github.com/cyverse-vice/}{https://github.com/cyverse-vice/} for \texttt{Dockerfile}s and configuration scripts; images are auto-built on DockerHub at \href{https://hub.docker.com/u/cyversevice}{https://hub.docker.com/u/cyversevice}.}.
These images include a reverse proxy using `nginx` to handle communication with CyVerse's authentication systems, which allow users to invite other registered users to securely access the same instance \citep{rstudio_proxy_2020}.
The images further include specific tools, such as the CLI utility `icommands` to access iRODS.
Users can reference the same base image and add packages to further customize their R environment, or they can use CyVerse's image registry where a user's published, i.e., publicly available, containers are held for additional reproducibility.
CyVerse user accounts are free (with some limitations for non-US users) and the [CyVerse Learning Center](https://learning.cyverse.org/) allows community members to document their containers to support training and education.

@@

@@

Nonetheless, at least on the level of R packages some **consolidation** seems in order, e.g., to reduce the number of packages creating `Dockerfile`s from R code or controlling the Docker daemon with R code.
It remains to be seen which approach to control Docker, via the Docker API as \pkg{stevedore} or via system calls as \pkg{dockyard}/\pkg{docker}/\pkg{dockr}, is more sustainable, or whether the question will be answered by the endurance of maintainers and sufficient funding.
Similarly, the capturing of environments and their serialisation in form of a `Dockerfile` currently happens at different levels of abstraction, and re-use of functionality seems reasonable, e.g., \pkg{liftr} could generate the environment with \pkg{containerit}, which in turn may use \pkg{dockerfiler} for low-level R objects representing a `Dockerfile` and its instructions.
In this consolidation, the Rocker&nbsp;Project could play the role of a coordinating entity.
Nonetheless, for the moment, it seems that Rocker Project will be used more experimentation and feature growth, e.g., images for GPU-based computing and artificial intelligence.
Even with coding being more and more accepted as a required and achievable skill, an easier access, for example by exposing containerisation benefits via simple user interfaces in the users' IDE, could be an important next step, since currently containerisation happens more in the background for UI-based development (e.g., a `rocker/rstudio` image in the cloud).
Furthermore, a maturing of the Rockerverse packages for managing containers may lead to them being adopted in situations where manual coding is currently required, e.g. in the case of \CRANpkg{RSelentium} or \CRANpkg{drake} (see Sections&nbsp;\nameref{rselenium} and \nameref{drake} respectively).
In some cases, e.g., for \CRANpkg{analogsea}, the interaction with the Docker daemon may remain too specific to re-use first-order packages to control Docker.
